---
title: "Creating Pretty Documents from R Markdown"
subtitle: "The Tactile Theme"
author: "Yixuan Qiu"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: vignette
    toc: true
toc-title: "Table of Contents"
vignette: >
  %\VignetteIndexEntry{Creating Pretty Documents from R Markdown - The Tactile Theme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dev.args = list(type = "cairo"))
```

> Have you ever tried to find a lightweight yet nice theme for the R Markdown
documents, just like this page?

## Themes for R Markdown

With the powerful `rmarkdown` package, we could easily create nice HTML document
by adding some meta information in the header, for example

```yaml
---
title: Nineteen Years Later
author: Harry Potter
date: July 31, 2016
output:
  rmarkdown::html_document:
    theme: lumen
---
```

The `html_document` engine uses the [Bootswatch](https://bootswatch.com/)
theme library to support different styles of the document.
This is a quick and easy way to tune the appearance of your document, yet with
the price of a large file size (> 700KB) since the whole
[Bootstrap](https://getbootstrap.com/) library needs to be packed in.

For package vignettes, we can use the `html_vignette` engine to generate
a more lightweight HTML file that is meant to minimize the package size, but
the output HTML is less stylish than the `html_document` ones.

So can we do **BOTH**, a lightweight yet nice-looking theme for R Markdown?

## The prettydoc Engine

The answer is YES! (At least towards that direction)

The `prettydoc` package provides an alternative engine, `html_pretty`,
to knit your R Markdown document into pretty HTML pages.
Its usage is extremely easy: simply replace the
`rmarkdown::html_document` or `rmarkdown::html_vignette` output engine by
`prettydoc::html_pretty` in your R Markdown header, and use one of the built-in
themes and syntax highlighters. For example

```yaml
---
title: Nineteen Years Later
author: Harry Potter
date: July 31, 2016
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
```

## Options and Themes

The options for the `html_pretty` engine are mostly compatible with the default
`html_document`
(see the [documentation](https://bookdown.org/yihui/rmarkdown/html-document.html))
with a few exceptions:

1. Currently the `theme` option can take the following values. More themes will
be added in the future. 
    - `cayman`: Modified from the [Cayman](https://github.com/jasonlong/cayman-theme) theme.
    - `tactile`: Modified from the [Tactile](https://github.com/jasonlong/tactile-theme) theme.
    - `architect`: Modified from the [Architect](https://github.com/jasonlong/architect-theme) theme.
    - `leonids`: Modified from the [Leonids](https://github.com/renyuanz/leonids) theme.
    - `hpstr`: Modified from the [HPSTR](https://github.com/mmistakes/jekyll-theme-hpstr) theme.
2. The `highlight` option takes value from `github` and `vignette`.
3. A new `math` parameter to choose between `mathjax` and `katex` for rendering math expressions.
   The `katex` option supports offline display when there is no internet connection.
4. Options `code_folding`, `code_download` and `toc_float` are not applicable.

## Offline Math Expressions

By default, `html_pretty` uses MathJax to render math expressions, for example inline math
expressions $x^2 + y^2 = z^2$, and display formulas:

$$
f(x)=\frac{1}{\sqrt{2\pi}\sigma}e^{-\frac{(x-\mu)^2}{2\sigma^2}}
$$

However, using MathJax requires an internet connection. If you need to create documents that can
show math expressions offline, simply add one line `math: katex` to the document metadata:

```yaml
---
title: Nineteen Years Later
author: Harry Potter
date: July 31, 2016
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    math: katex
---
```

This option will enable [KaTeX](https://katex.org/) for rendering the math expressions, and all
resource files will be included in for offline viewing. The offline document will be ~800k larger.

## Elements

We demonstrate some commonly used HTML elements here to show the appearance of themes.

### Headers

#### Level 4

##### Level 5


### Tables

|           | Df | Sum Sq | Mean Sq | F value | Pr(>F)  |    |
|-----------|----|--------|---------|---------|---------|----|
| Block     | 5  | 343.3  | 68.66   | 4.447   | 0.01594 | *  |
| N         | 1  | 189.3  | 189.28  | 12.259  | 0.00437 | ** |
| P         | 1  | 8.4    | 8.40    | 0.544   | 0.47490 |    |
| K         | 1  | 95.2   | 95.20   | 6.166   | 0.02880 | *  |
| N:P       | 1  | 21.3   | 21.28   | 1.378   | 0.26317 |    |
| N:K       | 1  | 33.1   | 33.14   | 2.146   | 0.16865 |    |
| P:K       | 1  | 0.5    | 0.48    | 0.031   | 0.86275 |    |
| Residuals | 12 | 185.3  | 15.44   |         |         |    |


### Lists

There are three kinds of lies:

1. Lies
2. Damned lies
3. Statistics
    - Frequentists
    - Bayesian
    - ...

Supported highlighters in `prettydoc`:

- `github`: Style similar to Github
- `vignette`: Style used by `rmarkdown::html_vignette`


### Markups

**Bold**, *italic*, <del>don't say this</del>.


### Code

Familiar `knitr` R code and plots:

```{r fig.width=6, fig.height=6, fig.align='center'}
set.seed(123)
n <- 1000
x1  <- matrix(rnorm(n), ncol = 2)
x2  <- matrix(rnorm(n, mean = 3, sd = 1.5), ncol = 2)
x   <- rbind(x1, x2)
smoothScatter(x, xlab = "x1", ylab = "x2")
head(x)
```

Also try some other languages, for example C++.

```cpp
// [[Rcpp::depends(RcppEigen)]]
// [[Rcpp::depends(RcppNumerical)]]

#include <RcppNumerical.h>

using namespace Numer;

// f = 100 * (x2 - x1^2)^2 + (1 - x1)^2
// True minimum: x1 = x2 = 1
class Rosenbrock: public MFuncGrad
{
public:
    double f_grad(Constvec& x, Refvec grad)
    {
        double t1 = x[1] - x[0] * x[0];
        double t2 = 1 - x[0];
        grad[0] = -400 * x[0] * t1 - 2 * t2;
        grad[1] = 200 * t1;
        return 100 * t1 * t1 + t2 * t2;
    }
};

// [[Rcpp::export]]
Rcpp::List optim_test()
{
    Eigen::VectorXd x(2);
    x[0] = -1.2;
    x[1] = 1;
    double fopt;
    Rosenbrock f;
    int res = optim_lbfgs(f, x, fopt);
    return Rcpp::List::create(
        Rcpp::Named("xopt") = x,
        Rcpp::Named("fopt") = fopt,
        Rcpp::Named("status") = res
    );
}
```
